% !TEX root = ../main.tex

\section{Entanglement and Joint Distributions}
\label{sec:entanglement}
\frame[plain]{\sectionpage}

%----------------------------------------
\begin{frame}{What entanglement provides}
    \textbf{Without entanglement} (single-qubit gates only): the
    state stays a product
    $|\psi\rangle = |\psi_1\rangle\otimes\cdots\otimes
    |\psi_{n_q}\rangle$ and measurements are
    \emph{statistically independent}:
    \[
    P(\mu_1,\ldots,\mu_{n_q}) = \prod_i P(\mu_i)
    \]
    \medskip

    \textbf{With CNOTs}: state is no longer separable. Joint
    structure that no product distribution can reproduce.
    \medskip

    \begin{block}{Bell state example}
        Two qubits, Hadamard on qubit 1, then CNOT:
        \[
        |0\rangle\otimes|0\rangle \;\longrightarrow\;
        \tfrac{1}{\sqrt 2}(|00\rangle+|11\rangle)
        \]
        Marginals: $\langle Z_1\rangle = \langle Z_2\rangle = 0$
        (uniform random).
        Joint: $\langle Z_1 Z_2\rangle = +1$ (perfectly
        correlated). No product of independent random variables
        can produce this.
    \end{block}
\end{frame}

%----------------------------------------
\begin{frame}{Two routes to correlated samples}
    \begin{columns}[T]
    \begin{column}{.48\textwidth}
        \textbf{Classical generator}\\
        Independent noise $\mathbf{z}$\\
        $\downarrow$ deterministic mixing\\
        $\downarrow$ via learned $W_1, W_2, W_3$\\
        $\downarrow$\\
        Correlated outputs
        \medskip

        All correlation must be encoded in the weight
        matrices.
    \end{column}
    \hfill
    \begin{column}{.48\textwidth}
        \textbf{Hybrid quantum generator}\\
        Independent noise $\mathbf{z}$\\
        $\downarrow$ entangled circuit\\
        $\downarrow$ produces correlated $(\mu_1,\ldots,\mu_{n_q})$\\
        $\downarrow$\\
        Classical head upsamples
        \medskip

        Quantum part introduces dependence between features
        \emph{before} the head sees them.
    \end{column}
    \end{columns}
    \medskip

    \begin{center}
    \emph{Hypothesis: this should make the hybrid generator
    \textbf{more parameter-efficient} for joint distributions ---
    a claim our project tests empirically.}
    \end{center}
\end{frame}

%----------------------------------------
\begin{frame}{What this is, and is not, a claim of}
    \begin{columns}[T]
    \begin{column}{.48\textwidth}
        \textbf{We claim, theoretically:}
        \begin{itemize}
            \item Entanglement provides an architectural
            \emph{pathway} for joint structure that
            classical mixing does not have natively
            \item The set of joint distributions reachable
            by depth-$L$ quantum circuits differs from the
            set reachable by classical NNs of comparable
            parameter count
        \end{itemize}
    \end{column}
    \hfill
    \begin{column}{.48\textwidth}
        \textbf{We do not claim:}
        \begin{itemize}
            \item That classical NNs cannot represent
            these distributions (they can, given enough
            capacity)
            \item That parameter-efficiency manifests in
            every regime
            \item That entanglement is the \emph{only}
            possible explanation for any observed effect
        \end{itemize}
    \end{column}
    \end{columns}
    \medskip

    \begin{center}
    \emph{The honest framing: a hypothesis we test, not a theorem
    we prove.}
    \end{center}
\end{frame}
